### 2023-08-22 resilience.tex
up

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
% Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

We now briefly investigate the question to control a system such that its controlled behavior may always reach $Safe$.
%
Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{Safe}$. 
%
%		\begin{definition}
We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient if $S  \xrightarrow{*} Safe$ is satisfied.  
%		\end{definition}
%
We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(Safe)$-resilient.

The \emph{maximal $Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{Safe}=(S,\rightarrow_{Safe},\leq)$ where $\rightarrow_{Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{Safe}$ is $Safe$-resilient.
%\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS and most of concrete WSTS allow such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

\begin{theorem}{}
Given an effective WSTS with effective pred-basis and an upward-closed set $Safe$, the maximal resilient subsystem is a computable WSTS.
\end{theorem}

\begin{proof}
Since $Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(Safe)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$. \qed
		\end{proof}
%
%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.

>>>>

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
. Since $\Safe \subseteq \pred^*(\Safe)$, one only need to decide whether 
the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.
In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
% Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

We now briefly investigate the question to control a system such that its controlled behavior may always reach $\Safe$.
%
Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $\Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{\Safe}$. 
%
%		\begin{definition}
We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient if $S  \xrightarrow{*} \Safe$ is satisfied.  
%		\end{definition}
%
We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(\Safe)$-resilient.

The \emph{maximal $\Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{\Safe}=(S,\rightarrow_{\Safe},\leq)$ where $\rightarrow_{\Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{\Safe}$ is $\Safe$-resilient.
%\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS and most of concrete WSTS allow such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

\begin{theorem}{}
Given an effective WSTS with effective pred-basis and an upward-closed set $\Safe$, the maximal resilient subsystem is a computable WSTS.
\end{theorem}

\begin{proof}
Since $\Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(\Safe)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(\Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
		\end{proof}
%
%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.
